---
title: "T-Test"
author: "Oleg Sechovcov"
date: "2024-10-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```




# Statudents-t distribution

dt(x, df)
pt(x, df)
qt(p, df)
rt(n, df)

```{r}
ggplot(data.frame(x = seq(-4,4, length=100)), aes(x=x)) +
  stat_function(fun = dt, args = list(df=4), color = 'orange')
```
# Hyothesis Testing
0) Check CLT conditions
1) Set-up hypothesis
2) Assume threshold values (alpha significance level)
3) Calculate
*   Point estimate
*   Standard error
*   Degree of freedom
*   t_statistic
*   p-value
4) Draw conclusions

# One sample t-test

```{r}
setwd("C:/Users/olegs/Documents/SDU - Copy/SDU/R/Data")
kiwi <- readr::read_csv('kiwi.csv')
```


We want to check if mean male GS kiwi height is 50 cm

0) Check CLT conditions
  * Data is independent - YES
  * Distribution is not strongly skewed - YES
  
```{r}
MGS <- kiwi %>% filter(Gender == 'M', Species_code == 'GS')
ggplot(MGS) +
  geom_histogram(aes(x=`Height(cm)`))
```


1) Set-up hypothesis
$H_0$ mean male GS kiwi height is equal to 50 cm $\mu = 50$
$H_A$ mean male GS kiwi height is not equal to 50 cm  $\mu \neq 50$

2) Assume threshold values (alpha significance level)

0.05

3) Calculate
*   Point estimate

```{r}
(pe_MGS <- mean(MGS$`Height(cm)`))
```
*   Standard error
```{r}
(sd_MGS <- sd(MGS$`Height(cm)`))
(SE_MGS <- sd_MGS/sqrt(nrow(MGS)))
```

*   Degree of freedom
```{r}
(df_MGS <- nrow(MGS)-1)
```

*   t_statistic
```{r}
(ts_MGS <- (pe_MGS - 50)/SE_MGS)
```
```{r}
ggplot(data.frame(x = seq(-4,4, length=100)), aes(x=x)) +
  stat_function(fun = dt, args = list(df= df_MGS), color = 'orange') +
  geom_vline(xintercept = ts_MGS)
```

*   p-value
```{r}
(pv_MGS <- 2 * pt(ts_MGS, df = df_MGS))
```
```{r}
pv_MGS < 0.05
```
4) Draw conclusions
We reject null hypothesis in favour of the alternative. The
mean male great spotted height is significantly different from assumed 50 cm

# One sample t-test short
```{r}
t.test(MGS$`Height(cm)`, MU=50)
```
# Difference of means t-test

kiwi: Is the average weight of female NIBr kiwi equal to the average weight of a
female Tok Kiwi

0) Check CLT conditions

  * Data is independent - YES
  * Distribution is not strongly skewed - YES
  
```{r}
FN <- kiwi %>% filter(Gender == 'F', Species_code == 'NIBr')
FT <- kiwi %>% filter(Gender == 'F', Species_code == 'Tok')
```

```{r}
ggplot(FN) +
  geom_histogram(aes(x=`Weight(kg)`))

ggplot(FT) +
  geom_histogram(aes(x=`Weight(kg)`))
```


1) Set-up hypothesis
$H_0$ The average weight of FNK is equal to average weight of a FTK $$\mu_{FNK}-\mu_{FTK}=0$$
$H_A$ The average weight of FNK is NOT equal to average weight of a FTK   $$\mu_{FNK}-\mu_{FTK} \neq 0$$

2) Assume threshold values (alpha significance level)

0.05

3) Calculate
*   Point estimate
```{r}
(pe_FK <- mean(FN$`Weight(kg)` - mean(FT$`Weight(kg)`)))
```
*   Standard error
$SE = \sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}$


```{r}
sd_FN <- sd(FN$`Weight(kg)`)
sd_FT <- sd(FT$`Weight(kg)`)

n_FN <- nrow(FN)
n_FT <- nrow(FT)

SE_FN <- sd_FN/sqrt(n_FN)
SE_FT <- sd_FT/sqrt(n_FT)

(SE_FK <- sqrt(SE_FN^2 + SE_FT^2))

```
*   Degree of freedom
```{r}
(df_FN <- nrow(FN) - 1)
(df_FT <- nrow(FT) - 1)

(df_FK <- min(df_FN,df_FT))
```
*   t_statistic
```{r}
(ts_FK <- (pe_FK - 0)/SE_FK)
```
```{r}
ggplot(data.frame(x = seq(-4,4, length=100)), aes(x=x))+
  stat_function(fun=dt, args = list(df = df_FK))+
  geom_vline(xintercept = ts_FK, color='red')
```

*   p-value
```{r}
(pv_FK <- 2* (1 -pt(ts_FK, df=df_FK)))
```

4) Draw conclusions

We keep null hypothesis. There is no significant difference between average female
NIBr weight and average female Tok weight

# Difference of means t-test Short

```{r}
t.test(FN$`Weight(kg)`, FT$`Weight(kg)`)
```


# Paired data t-test

```{r}
setwd("C:/Users/olegs/Documents/SDU - Copy/SDU/R/Data")
ap <- readr::read_delim('antiper1.dat', delim = ';')
```

```{r}
nrow(ap)
```
```{r}
head(ap,5)
```

```{r}
ggplot(ap)+
  geom_density(aes(x= DryPowderScore), fill='purple', alpha = 0.3)+
  geom_density(aes(x= PowderInOilScore), fill='green', alpha = 0.3)
```
```{r}
ap <- ap %>% mutate(diff = PowderInOilScore - DryPowderScore)
```

```{r}
ggplot(ap)+
  geom_density(aes(x=diff))
```
0) Check CLT conditions

  * Data is independent - YES
  * Distribution is not strongly skewed - YES

1) Set-up hypothesis

$H_0$ The average difference in scoring between PowderInOil and DrPowder is equal to 0 $$\mu_{diff}=0$$
$H_A$ The average difference in scoring between PowderInOil and DrPowder is NOT equal to 0   $$\mu_{diff} \neq 0$$

2) Assume threshold values (alpha significance level)

0.05

3) Calculate
*   Point estimate

```{r}
(pe_ap <- mean(ap$diff))
```
*   Standard error
```{r}
sd_ap <- sd(ap$diff)
n_ap <- nrow(ap)
(SE_ap <- sd_ap/sqrt(n_ap))

```
*   Degree of freedom
```{r}
(df_ap <- n_ap -1)
```
*   t_statistic
```{r}
(ts_ap <- (pe_ap - 0)/SE_ap)
```
```{r}
ggplot(data.frame(x=seq(-4,4, lenght=100)), aes(x=x))+
  stat_function(fun=dt, args = list(df=df_ap))+
  geom_vline(xintercept = ts_ap, color='red')
```

*   p-value
```{r}
(pv_ap <- 2* pt(ts_ap, df= df_ap))

```
4) Draw conclusions
We reject null hypothesis in favour of alternative. There is a statistically significant difference 
between two types of antiperspirants


# Paired data t-test Short

```{r}
t.test(ap$diff, mu=0)
```


