---
title: "Data Visualization"
author: "E.Pastucha"
date: "2024-07-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Data sets

IRIS, MPG

```{r}
head(iris, 5)
nrow(iris)
```

```{r}
head(mpg, 5)
nrow(mpg)
```

# Scatter plot

```{r}
ggplot(data = iris) + 
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, color = Species)) + 
  labs(x = 'Petal Length [cm]',
       y = 'Petal Width [cm]',
       color = 'Iris Species',
       title = 'Iris flowers - Petal Length vs Petal Width', 
       subtitle = 'Colored by Species')
```

```{r}
ggplot(data = iris) + 
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, 
                           #shape = Species,
                           #size = Species
                           color = Species)) + 
  labs(x = 'Petal Length [cm]',
       y = 'Petal Width [cm]',
       size = 'Iris Species',
       title = 'Iris flowers - Petal Length vs Petal Width', 
       subtitle = 'Colored by Species') + 
  coord_fixed(ratio = 1)
```

```{r}
ggplot(data = iris) + 
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, 
                           #shape = Species,
                           #size = Species
                           color = Species)) + 
  labs(x = 'Petal Length [cm]',
       y = 'Petal Width [cm]',
       size = 'Iris Species',
       title = 'Iris flowers - Petal Length vs Petal Width', 
       subtitle = 'Colored by Species') + 
  coord_fixed(ratio = 1) + 
  geom_smooth(mapping = aes(x = Petal.Length, y = Petal.Width), method = lm)
```

```{r}
ggplot(data = iris) + 
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, 
                           color = Species)) + 
  geom_smooth(mapping = aes(x = Petal.Length, y = Petal.Width), method = lm)+ 
  geom_smooth(mapping = aes(x = Petal.Length, y = Petal.Width, color = Species), method = lm)
```

# Facet wrap and facet grid

```{r}
ggplot(data = iris) + 
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width)) +
  facet_wrap(~Species)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = displ)) + 
  facet_grid(class~factor(cyl))
```

# Naming


```{r}
ggplot(iris) + 
  geom_point(aes(Petal.Length, Petal.Width, 
                           #shape = Species,
                           #size = Species
                           color = Species)) + 
  labs(x = 'Petal Length [cm]',
       y = 'Petal Width [cm]',
       size = 'Iris Species',
       title = 'Iris flowers - Petal Length vs Petal Width', 
       subtitle = 'Colored by Species') + 
  coord_fixed(ratio = 1)
```

# Box Plots

```{r}
ggplot(iris) + 
  geom_boxplot(mapping = aes(y = Species,
                             x = Petal.Length,
                             fill = Species),
               alpha = 0.2,
               color = 'darkred') + 
  labs(x = 'Petal Length [cm]', 
       y = 'Iris Species')
```

```{r}
ggplot(mpg) + 
  geom_boxplot(aes(x = class, y = hwy, color = class), show.legend = FALSE)
```

# Histogram and Density Plot

```{r}
ggplot(iris) + 
  geom_histogram(aes(x = Petal.Length, fill = Species), binwidth = 0.2, position = 'dodge')
```

```{r}
versicolor <- filter(iris, Species == 'versicolor')
```

```{r}
ggplot(versicolor) + 
  geom_histogram(aes(x = Petal.Length), color = 'black', fill ='limegreen', binwidth = 0.1)
```
```{r}
ggplot(versicolor) + 
  geom_density(aes(x = Petal.Length), fill = 'yellow', 
               alpha = 0.3, 
               color = 'red')
```

```{r}
ggplot(versicolor) + 
  geom_histogram(aes(x = Petal.Length, y = after_stat(density)), 
                 fill = 'limegreen', binwidth = 0.1)+
  geom_density(aes(x = Petal.Length), color = 'red', fill = 'yellow', alpha = 0.3)
```

```{r}
ggplot(iris) + 
  geom_density(aes(x = Petal.Length, fill = Species), alpha = 0.3)
```
```{r}
ggplot(iris) + 
  geom_density(aes(x = Petal.Length, fill = Species), alpha = 0.3) + 
  geom_histogram(aes(x = Petal.Length, y = after_stat(density), fill = Species), 
                 alpha = 0.3)
```

# Empirical cumulative distribution function - stat_ecdf

```{r}
ggplot(iris) + 
  stat_ecdf(mapping = aes(x = Sepal.Length, color = Species))
```

# Barplots

```{r}
ggplot(mpg) + 
  geom_bar(mapping = aes(x = class, fill = manufacturer), color = 'black')
```

```{r}
library(viridis)
```



```{r}
ggplot(mpg) + 
  geom_bar(mapping = aes(x = class, fill = manufacturer), color = 'black') + 
  scale_fill_viridis(discrete = TRUE)
```

```{r}
ggplot(mpg) + 
  geom_bar(mapping = aes(x = manufacturer, fill = class), color = 'black') + 
  scale_fill_brewer(palette = 'YlOrBr') + 
  coord_flip()
```

BrBG, PiYG, PRGn, PuOr, RdBu, RdGy, RdYlBu, RdYlGn, Spectral
Accent, Dark2, Paired, Pastel1, Pastel2, Set1, Set2, Set3
Blues, BuGn, BuPu, GnBu, Greens, Greys, Oranges, 
OrRd, PuBu, PuBuGn, PuRd, Purples, RdPu, Reds, YlGn, YlGnBu, YlOrBr, YlOrRd

































