---
title: "Inference for Categorical Data"
author: "Oleg Sechovcov"
date: "2024-10-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```
# One sample proportion test

$$SE = \sqrt{\frac{p \cdot(1-p)}{n}}$$
0) Check CLT conditions.
1) Setup hypotheses
2) Assume threshold for values
* $\alpha = 0.05$ is common
3) Calculate the results
* point estimate: $\hat{p}$
* standard error: $SE = \sqrt{\frac{p \cdot(1-p)}{n}}$
* p-value: $2 \cdot P(Z > |z|)$
4) Draw a conclusion

$H_0$: 50% of NBA players is over 200 cm tall: $p = 0.5$

$H_a$: More than 50% of NBA players is over 200 cm tall: $p > 0.5$

```{r}
alpha <- 0.05
```

```{r}
setwd("C:/Users/olegs/Documents/SDU - Copy/SDU/R/Data")
nba <- readr::read_csv('nba_ht_wt.csv')
```

```{r}
nba <- nba %>% mutate(Height = Height* 2.54)
```

```{r}
n_success <- nba %>% filter(Height > 200) %>% nrow()

```
```{r}
nba %>% filter(Height <= 200) %>% nrow()
```
```{r}
n_sample <- nrow(nba)
(point_estimate_nba <-  n_success / n_sample)
```
```{r}
(SE_nba <- sqrt(0.5 * 0.5/n_sample))
```
```{r}
ggplot(data.frame(x = seq(0.4,0.6, length = 100)), aes(x=x)) +
  stat_function(fun = dnorm, args = list(mean = 0.5, sd = SE_nba)) +
  geom_vline(xintercept = point_estimate_nba, color = "red") 
```

```{r}
(p_value_nba <- (1-pnorm(point_estimate_nba, mean = 0.5, sd = SE_nba)))
```
We reject null hypothesis, because p-value is less than 0.05.


# Difference of proportions test

0) Check CLT conditions.
1) Setup hypotheses
2) Assume threshold for values
* $\alpha = 0.05$ is common
3) Calculate the results
* point estimate: $\hat{p}$
* standard error: $SE = \sqrt{\frac{p \cdot(1-p)}{n}}$
* p-value: $2 \cdot P(Z > |z|)$
4) Draw a conclusion

A clothes producer is looking for a new supplier of zipppers.Two factories are
frontrunners. The producer wants to decide based on one days production results.

First factory produces 23 935 zippers, out of which 132 were faulty.
Second factory produces 22 312 zippers, out of which 111 were faulty.
We want to check if it's safe to assume that they have the same production of faulty
zippers, and due to that we can choosed based squerly on price

$H_0$: Proportion of faulty zippers is the same in both factories: $p_1 = p_2$

$H_a$: Proportion of faulty zippers is different in both factories: $p_1 \neq p_2$


Pooled proportion: $\hat{p} = \frac{x_1 + x_2}{n_1 + n_2}$
```{r}
(p_pooled <- (132 + 111) / (23935 + 22312))
p_pooled * 23935
p_pooled * 22312
(1- p_pooled) * 23935
(1- p_pooled) * 22312
```
```{r}
alpha <- 0.05
```

```{r}
proportion1 <- 132 / 23935
proportion2 <- 111 / 22312
(point_estimate_zippers <- proportion1 - proportion2)
```
$$SE = \sqrt{\hat{p} \cdot(1-\hat{p}) \cdot \left(\frac{1}{n_1} + \frac{1}{n_2}\right)}$$
```{r}
(SE_zippers <- sqrt(p_pooled * (1-p_pooled) * (1/23935 + 1/22312)))
```
```{r}
ggplot(data.frame(x = seq(-0.0025,0.0025, length = 100)), aes(x=x)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = SE_zippers)) +
  geom_vline(xintercept = point_estimate_zippers, color = "red") 
```

```{r}
(p_value_zippers <- 2 * (1 - pnorm(point_estimate_zippers, mean = 0, sd = SE_zippers)))
```
```{r}
p_value_zippers > alpha
```
We accept Null hypothesis, because p-value is greater than 0.05. There is no evidence that the proportion of faulty zippers is different in both factories.


# $\chi^2$ distribution

```{r}
ggplot(data.frame(x=seq(0,30,length=100)), aes(x=x))+
  stat_function(fun = dchisq, args = list(df = 2), color = 'blue')+
  stat_function(fun = dchisq, args = list(df = 3), color = 'green')+
  stat_function(fun = dchisq, args = list(df = 5), color = 'yellow')+
    stat_function(fun = dchisq, args = list(df = 7), color = 'orange')+
      stat_function(fun = dchisq, args = list(df = 11), color = 'red')+
        stat_function(fun = dchisq, args = list(df = 19), color = 'purple')




```


# $\chi^2$ goodness of fit test

0) Check conditions.
1) Setup hypotheses
2) Setup threshold 
* $\alpha = 0.05$ is common
3) Calculate chi^2 statistic
4) Calculate p-value
5) Draw a conclusion

Sample of 669 Haribo Gummy Bears
```{r}
(gummy_bears <- c(83, 142, 100, 103, 104, 137))
```
$H_0$: The distribution of flavors is equal

$H_a$: The distribution of flavors is not equal

```{r}
alpha <- 0.05
```

```{r}
(expected <- sum(gummy_bears) / length(gummy_bears))
```
$$Z_n = \sum_{i=1}^{k} \frac{(O_i - E_i)^2}{E_i}$$
$$X^2 = \sum_{i=1}^{k} \frac{(O_i - E_i)^2}{E_i}$$
```{r}
(gummy_Z <- (gummy_bears - expected) / sqrt(expected))
```

```{r}
(chi_score <- sum(gummy_Z^2))
```
degrees of freedom: df = 5

```{r}
ggplot(data.frame(x=seq(0,30,length=100)), aes(x=x))+
  stat_function(fun = dchisq, args = list(df = 5), color = 'blue')+
  geom_vline(xintercept = chi_score, color = "red")
```
```{r}
(p_value <- 1 - pchisq(chi_score, df = 5))
```
```{r}
p_value > alpha
```
We reject null hypothesis, because p-value is less than 0.05. There is evidence that the distribution of flavors is not equal.


## Just run the test with no calculations

```{r}
chisq.test(x= gummy_bears, p = rep(c(expected), length(gummy_bears)), rescale.p = TRUE)
```
# $\chi^2$ test for Independence

```{r}
(movies_pizza <- tribble(
  ~Movie, ~Pepperoni, ~Mushrooms, ~Kebab,
  "Frozen 2", 20, 10, 5,
  "Joker", 15, 12, 15,
  "Someone Great", 8, 13,2
))

```
0) Check conditions.
1) Setup hypotheses
2) Get the observed count for each category
3) Calculate expected count for each category
4) Calculate test statistic
6) Calculate p-value
5) Draw a conclusion

$H_0$: There is no association between movie and pizza topping

$H_a$: There is an association between movie and pizza topping

```{r}
p_total <- sum(movies_pizza$Pepperoni)
m_total <- sum(movies_pizza$Mushrooms)
k_total <- sum(movies_pizza$Kebab)
total <- p_total + m_total + k_total
```

```{r}
(movies_pizza <- movies_pizza %>% 
        mutate(P_exp = (Pepperoni + Mushrooms + Kebab) * p_total/total)  %>% 
        mutate(M_exp = (Pepperoni + Mushrooms + Kebab) * m_total/total) %>% 
        mutate (K_exp = (Pepperoni + Mushrooms + Kebab) * k_total/total))
```

```{r}
(movies_pizza <- movies_pizza %>% 
  mutate(P_Z = ((Pepperoni - P_exp) / sqrt(P_exp))^2) %>% 
  mutate(M_Z = ((Mushrooms - M_exp) / sqrt(M_exp))^2) %>%
  mutate(K_Z = ((Kebab - K_exp) / sqrt(K_exp))^2))
 
```
```{r}
(chi_score_mp <- sum(movies_pizza$P_Z) + sum(movies_pizza$M_Z) + sum(movies_pizza$K_Z))
```
df = 4

```{r}
ggplot(data.frame(x=seq(0,15,length=100)), aes(x=x))+
  stat_function(fun = dchisq, args = list(df = 4), color = 'blue')+
  geom_vline(xintercept = chi_score_mp, color = "red")
```
```{r}
alpha <- 0.05

```

```{r}
(p_value_mp <- 1 - pchisq(chi_score_mp, df = 4))
```
```{r}
p_value_mp > alpha
```
We reject null hypothesis, because p-value is less than 0.05. There is evidence that there is an association between movie and pizza topping.



## Just run the test with no calculations


```{r}
(movies_pizza <- tribble(
  ~Movie, ~Pepperoni, ~Mushrooms, ~Kebab,
  "Frozen 2", 20, 10, 5,
  "Joker", 15, 12, 15,
  "Someone Great", 8, 13,2
))

```


```{r}
movies_pizza %>% select(-Movie) %>% chisq.test()
```


