---
title: "Data Transformation"
author: "Oleg Sechovcov"
date: "2024-09-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```
# Filter
C:\Users\olegs\Documents\SDU - Copy\SDU\R\Data

```{r}
setwd("C:/Users/olegs/Documents/SDU - Copy/SDU/R/Data")
```

```{r}
kiwi <- readr::read_csv('kiwi.csv')
```

```{r}
ncol(kiwi)
nrow(kiwi)
colnames(kiwi)
```
```{r}
filter(kiwi, Species_code == 'GS', Gender == 'M', `Weight(kg)` > 2.2)
```

[Ctr] + [Shift] + [M]
```{r}
%>% 
```

```{r}
kiwi %>% filter(Species_code == 'GS', Gender == 'M', `Weight(kg)` > 2.2)
```
```{r}
kiwi %>% filter(Species_code == 'GS') %>% 
  filter(Gender == 'M') %>% 
  filter(`Weight(kg)` > 2.2)
```

```{r}
kiwi %>% filter(Species_code != 'GS')

```
```{r}
kiwi %>% filter(Species_code != 'Tok' | Species_code == 'NIBr')
```
Logical opperators
& 
| 
>
>=
<
<=
!=
==

# Select()

```{r}
head(kiwi, 5)
```
```{r}
kiwi %>%  select(1:4)

```
```{r}
kiwi %>% select(Species_code, Gender)
```
```{r}
kiwi %>%  select(-Location)
```

# Mutate()

BMI = Weight(kg)/Height(m)^2 

```{r}
Kiwi_BMI <- kiwi %>%  
  mutate(`Height(m)` = `Height(cm)` / 100) %>% 
  mutate(BMI = `Weight(kg)` / `Height(m)`^2)

Kiwi_BMI
```

# summarise(), arrange(), group_by()
```{r}
kiwi %>% 
  group_by(Species_code) %>% 
  tally()
```
```{r}
kiwi %>% 
  group_by(Species_code, Gender) %>% 
  summarise(`mean weight` = mean(`Weight(kg)`), `mean height` = 
              mean(`Height(cm)`)) %>% 
  arrange(desc(`mean weight`)) %>% 
  knitr::kable()
```
# Altogehter now

```{r}
library(gapminder)
```

```{r}
head(gapminder, 5)
```

```{r}
unique(gapminder$year)
```

Biggest country (pop) in 1972:

```{r}
gapminder %>% 
  filter(year == 1972) %>% 
  arrange(desc(pop)) %>% 
  head(1) %>% 
  select(1,2,5) %>% 
  knitr::kable()
```

5 biggest countries in Africa 1987

```{r}
gapminder %>% 
  filter(year == 1987) %>% 
  filter(continent  == 'Africa') %>% 
  top_n(5, pop) %>% 
  knitr::kable()
```

dplyr + ggplot


Bar plot of 7 biggest countries in Americas in 1992
```{r}
 gapminder %>% 
  filter(continent == 'Americas') %>% 
  filter(year == '1992') %>% 
  arrange(desc(pop)) %>% 
  head(7) %>% 
  ggplot() +
  geom_bar(mapping = aes(x = reorder(country, pop), country, y = pop, fill = country), stat = 'identity')
  
```

```{r}
kiwi %>% 
  filter(Species_code == 'GS') %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = Gender, y = `Weight(kg)`))
```

```{r}
kiwi %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = Species_code, y = `Weight(kg)`, colour = Gender))
```

# Missing Values

```{r}
svalbard <- readr::read_csv('svalbard-climate-1912-2017.csv')
```

```{r}
svalbard %>% 
  ggplot() +
  geom_line(aes(x = YEAR, y = FEB), colour = 'blue') +
  geom_line(aes(x = YEAR, y = AUG), colour = 'red') 
```


```{r}
svalbard <- readr::read_csv('svalbard-climate-1912-2017.csv', na = c('NA','999.9'))
```

```{r}
svalbard %>% 
  ggplot() +
  geom_line(aes(x = YEAR, y = FEB), colour = 'blue') +
  geom_line(aes(x = YEAR, y = AUG), colour = 'red') 
```
```{r}
mean(svalbard$FEB, na.rm = TRUE)
sd(svalbard$JAN, na.rm = TRUE)
median(svalbard$APR, na.rm = TRUE)
```
```{r}
svalbard %>% 
  filter(!is.na(JAN), YEAR >= 2000)
```




