---
title: "Basic of inference - implementation"
author: "E.Pastucha"
date: "2024-08-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Normal distribution

dnorm(x, mean, sd) - distribution function
pnorm(x, mean, sd) - probability, 
qnorm(probability, mean, sd) - given probability what is the value
rnorm(how_many, mean, sd) - generate data 

default values: mean = 0, sd = 1

```{r}
ggplot(data.frame(x = seq(-10, 10, length = 100)), aes(x = x))+
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1), color = 'blue') + 
  stat_function(fun = dnorm, args = list(mean = 0, sd = 2), color = 'green') + 
  stat_function(fun = dnorm, args = list(mean = 0, sd = 3), color = 'yellow') +
  stat_function(fun = dnorm, args = list(mean = 2, sd = 1), color = 'orange')
```

```{r}
ggplot(data.frame(x = seq(-5, 5, length = 100)), aes(x = x))+
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1), color = 'blue') + 
  geom_vline(xintercept = 1, color ='red')
```

```{r}
ggplot(data.frame(x = seq(-5, 5, length = 100)), aes(x = x))+
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1), color = 'blue') + 
  geom_segment(aes(x = 1, y = 0, 
               xend = 1, yend = dnorm(1)), color ='red', linetype = 'dashed')
```

# Sharks

```{r}
sharks <- readr::read_csv('sharks.csv')
```
```{r}
head(sharks, 5)
```
```{r}
ggplot(sharks) + 
  geom_point(aes(x = `Total Length`, y = `Fork Length`, color = `Fish Sex`))
```

```{r}
ggplot(sharks) + 
  geom_density(aes(x = `Total Length`, fill = `Fish Sex`), alpha = 0.3)
```

```{r}
females <- filter(sharks, `Fish Sex` == 'Female')
```

```{r}
mean_fs <- mean(females$`Total Length`)
sd_fs <- sd(females$`Total Length`)
ggplot(females) + 
  geom_histogram(aes(x = `Total Length`, y = after_stat(density)), 
                 fill = 'red', alpha = 0.3, binwidth = 1) + 
  stat_function(fun = dnorm, args = list(mean = mean_fs, sd = sd_fs), size = 1)
```
# ECDF

```{r}
ggplot(females) + 
  stat_ecdf(aes(x = `Total Length`), color = 'limegreen') + 
  geom_line(stat = 'function', fun = pnorm, args = list(mean = mean_fs, sd = sd_fs), 
            color = 'purple')
```

Q-Q plot 

```{r}
ggplot(females) + 
  stat_qq_line(aes(sample = `Total Length`), color = 'red', size = 1) + 
  stat_qq(aes(sample = `Total Length`))
```

# Z-scores

150 inches long female shark.


```{r}
(z_score <- (150 - mean_fs)/sd_fs)
```

```{r}
pnorm(z_score, mean = 0, sd = 1)
pnorm(z_score)
```

```{r}
pnorm(150, mean = mean_fs, sd = sd_fs)
```

```{r}
ggplot(data.frame(x = seq(40, 220, length = 500)), aes(x=x)) + 
  stat_function(fun = dnorm, args = list(mean = mean_fs, sd = sd_fs)) + 
  geom_segment(aes(x = 150, y = 0,
                   xend = 150, yend = dnorm(150, mean = mean_fs, sd = sd_fs)), color = 'red') + 
  geom_area(stat = 'function', fun = dnorm, args = list(mean = mean_fs, sd = sd_fs), 
            fill = 'blue', xlim = c (40, 150), alpha = 0.3)
```

# SE

Mean female shark Total Length. 
1155 cases, 40-220 inches. 

```{r}
mean_fs
```

Central Limit Theorem:
* Samples are independent
* Sample size is bigger than 30 
* Population distribution is not strongly skewed

YES!

$$SE = \frac{\sigma}{\sqrt{n}}$$
```{r}
(SE <- sd_fs/sqrt(nrow(females)))
```

95% confidence interval.

1.96 
qnorm()

```{r}
mean_fs - 1.96 *SE
mean_fs + qnorm(0.025)*SE
```

```{r}
mean_fs + 1.96*SE
mean_fs + qnorm(0.975)*SE
```

We are 95% confident that population mean female shark total length is in between 115.54 inch
and 118.18 inch.

99% confidence interval

```{r}
mean_fs - 2.58 *SE
mean_fs + qnorm(0.005)*SE
```

```{r}
mean_fs + 2.58*SE
mean_fs + qnorm(0.995)*SE
```

We are 99% confident that population mean total length of a female shark is between 115.12 inch and 118.6 inch.

# Distribution of a sample - one sample (ONE!)

```{r}
ggplot(data.frame(x = seq(40, 220, length = 500)), aes(x = x)) + 
  stat_function(fun = dnorm, args = list(mean = mean_fs, sd = sd_fs))+
labs(title = 'Distribution of a sample of female sharks ')
```

# Sampling Distribution - multiple samples.


```{r}
ggplot(data.frame(x = seq(40, 220, length = 500)), aes(x = x)) + 
  stat_function(fun = dnorm, args = list(mean = mean_fs, sd = SE))+
labs(title = 'Distribution of a sample of female sharks ')
```


```{r}
ggplot(data.frame(x = seq(40, 220, length = 500)), aes(x = x)) + 
  stat_function(fun = dnorm, args = list(mean = mean_fs, sd = sd_fs), color = 'red')+
  stat_function(fun = dnorm, args = list(mean = mean_fs, sd = SE), color = 'green')+
  labs(title = 'Distribution of a sample of female sharks ')
```

































