---
title: "2022 01 03 VB-STA5 Exam in Statistics"
author: "Oleg Sechovcov"
date: "2025-01-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. North America Rodents
Dataset data/surveys.csv contains information about rodents sightings in North America from 1977 to 2002.
Dataset data/species.csv contains information about species acronyms and their Genus.

##a) Join the two datasets

```{r}
library(tidyverse)
Surveys <- readr::read_csv('data/surveys.csv')
Species <- readr::read_csv('data/species.csv ')
```
```{r}
Rodents <- Surveys %>% left_join(Species, by ='species_id')
head(Rodents,5)
```
##b) Present the 5 rodent species having the highest mean weight in a table showing species, mean weight and mean hindfoot length as in the example below. 
```{r}
Rodents %>% 
  group_by(species) %>% 
  summarise(`Mean Weight [g]` = mean(weight), `Mean Hindfoot Length [mm]` = mean(hindfoot_length)) %>% 
  arrange(desc(`Mean Weight [g]`)) %>% 
  head(5)
```
##c) Recreate the plot
```{r}
Rodents %>% 
  filter(year >= 1980 & year <= 1989) %>% 
  filter(genus == 'Dipodomys') %>% 
ggplot()+
  geom_point(aes(y = weight, x= hindfoot_length, color = species))+
  labs(x='Hindfoot Length [mm]', y='Weight [g]', title='Dipodomys - Kongaroo Rats', subtitle = 'caught in years 1980-1989' )
```
##d) Describe the plot
• The plot presents relationship between Kangooro Rats hindfoot length and their weight.
• There are 3 species within this genom. Merriami, ordii and spectabilis.
• Spectabilis species is heavier and has bigger feet than the other species.
• Ordii seems slightly heavier than merriami.
• There are multiple outliers - maybe due to mislabeling.
• There seems to be linear relationship in between 2 variables, but not very strong.

##e) Kangaroo Rats (genus Dipodomys) are small rodents moving similarily to kangaroos using jumping

We will be conducting a difference of mean t-test

H_0 There is no significant mean difference in hindfoot length between ordii and merriami species
H_1 There is a significant mean difference in hindfoot length between ordii and merriami species

```{r}
Kangaroo_M <- Rodents %>% 
  filter(sex == 'M') %>% 
  filter(species == 'merriami')

ggplot(Kangaroo_M)+
  geom_histogram(aes(x = hindfoot_length))

Kangorro_O <- Rodents %>% 
   filter(sex == 'M') %>% 
  filter(species == 'ordii')

ggplot(Kangorro_O)+
  geom_histogram(aes(x = hindfoot_length))
```

```{r}
Rodents %>% 
  filter(species == 'ordii' | species == 'merriami') %>% 
  filter(sex == 'M') %>% 
  t.test(hindfoot_length~species, data = .)
```
#2. Medical students smoking habits

A study was conducted on various Medical Universities within Germany and Hungary. The students were
asked about their smoking habits. 2883 students took part, 44% of them were German, 36% were Hungarian
and 20% other nationalities. The table below lists number of students per nationality declaring daily smoking
habit.

##a) Are the proportions of nationalities within smoking students a true representation of proportions of the whole student body? Conduct a suitable test to check this hypothesis


H_0 The proportions of nationalities are within the smoking student representation of proportions of the whole student body

H_1 The proportions of nationalities are not within the smoking student representation of proportions of the whole student body

## 44%
```{r}
all_smoking <- 91 + 78 + 51
(expected_G <- 0.44 * all_smoking)
```
## 36%
```{r}
all_smoking <- 91 + 78 + 51
(expected_G <- 0.36 * all_smoking)
```


## 10%
```{r}
all_smoking <- 91 + 78 + 51
(expected_G <- 0.20 * all_smoking)
```
## Chi-squared
```{r}
chisq.test(c(91,78,51), p=c(0.44,0.36,0.20))
```
#3. University salaries
```{r}
Salaries <- readr::read_csv('data/salaries.csv')
colnames(Salaries)
```
"sx" - Possible important
"rk" - important
"yr" - Possible important 
"dg" - important 
"yd" - important
"sl" - to predict

```{r}
ggplot(Salaries)+
  geom_point(aes(x = rk, y = sl))
  
```
## Rˆ2 Backward Elimination
```{r}
fit <- lm(sl ~ #sx +
            rk+
            yr
            #dg+
            #yd
            ,data = Salaries)
summary(fit)
```
### Iteration 1
All included 0.8357 R^2

```{r}
plot(fit)
```
```{r}
(predicted_salery <- 16203.27 + 4262.28 + 375.70 * 10)
```


