---
title: "2024 01 08 VB-STA5 Exam in statistics - Solution Guide"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}
suppressPackageStartupMessages(library(tidyverse))
```

# 1. Math and Statistics exam performance

Dataset *data/students_exam_performance.csv* contains information about students that participated both in Mathematics and Statistics class.

a) Recreate the plot:

```{r}
students <- readr::read_csv('data/students_exam_performance.csv')

ggplot(students) + 
  geom_boxplot(aes(y = major, x = statistics_score))+
  labs(x = 'Statistics Exam Scores',
       y = 'Major',
       title = 'Results of Statistics exam',
       subtitle = 'Divided according to Major') 
```


b) Describe the plot.

* boxplot presenting statistics exam score according to the students major

* mean score per major are in order (from lowest to highest) Social Sciences, Political Science, Economics and Finance, Environmental Sciences, Biology and last Mathematics and Statistics.

* social sciences lowest score ranging from 30-60 points.

* ... for the rest of majors

* Mathematics and Statistics highest score ranging from 75 to 97-8. Also with narrowest distribution. Except for few outliers students scored in 80-100 points range. 


c) Check whether there is a significant difference between a Mathematics Exam score for *Economics and Finance* major students with minor in *Mathematics and Statistics*, and *Economics and Finance* major students with other minors. Conduct an appropriate test for this situation.    


Difference of means t-test.

$H_0:\mu_{m\_minor_ms} - \mu_{m\_minor_not_ms}=0$

$H_A:\mu_{m\_minor_ms} - \mu_{m\_minor_not_ms}=0 \neq0$

H0: There is no difference between mean Math exam score for *Economics and Finance* major students with minor in *Mathematics and Statistics*, and *Economics and Finance* major students with other minors

HA: There is difference between mean Math exam score for *Economics and Finance* major students with minor in *Mathematics and Statistics*, and *Economics and Finance* major students with other minors

alpha significance level - 0.05

onditions check:

Normality:

```{r}
students_b_m <- students %>% 
  filter(major == 'Economics and Finance') %>% 
  filter(minor == 'Mathematics and Statistics')
students_b_nm <- students %>% 
  filter(major == 'Economics and Finance') %>% 
  filter(minor != 'Mathematics and Statistics')

ggplot(students_b_m) + 
  geom_histogram(aes(x = math_score))
ggplot(students_b_nm) + 
  geom_histogram(aes(x = math_score))
```

The variables distributions look normal.

We assume that observations are independent.

* short version

```{r}
t.test(students_b_m$math_score, students_b_nm$math_score)
```

p-value is smaller than alpha significance level, thus we reject null hypothesis in favour of the alternative. There is statistically significant difference between mean Math exam score for *Economics and Finance* major students with minor in *Mathematics and Statistics*, and *Economics and Finance* major students with other minors

* long version

```{r}
(point_estimate <- mean(students_b_m$math_score) - 
   mean(students_b_nm$math_score))
(nrow(students_b_m))
(nrow(students_b_nm))
```

```{r}
dof <- 152
```

```{r}
(SE <- sqrt((sd(students_b_m$math_score)^2/nrow(students_b_m)) + 
              (sd(students_b_nm$math_score)^2/nrow(students_b_nm))))
```

```{r}
(t_score <- (point_estimate - 0)/SE)
```

```{r}
ggplot(data.frame(x = seq(-5, 5, length=100)), aes(x = x)) +
  stat_function(fun = dt, args = list(df = dof), color = 'blue') +
  geom_vline(aes(xintercept = t_score),  color = 'red')
```

```{r}
(p_value <- 2 * (1- pt(t_score, df = dof)))
```
p-value is smaller than alpha significance level, thus we reject null hypothesis in favour of the alternative. There is statistically significant difference between mean Math exam score for *Economics and Finance* major students with minor in *Mathematics and Statistics*, and *Economics and Finance* major students with other minors


# Childs seatbelt - car seat legislation

```{r}
accidents <- readr::read_csv("data/CarSeatLegislation.csv")
```


a) Present the number and proportion of 'No Injury' accidents divided according to the implemented protection.

```{r}
accidents %>% 
  filter(Injury == 'No Injury') %>% 
  group_by(Restraint) %>% 
  tally() %>% 
  mutate(Proportion = n/sum(n)) %>% 
  knitr::kable()
```

b) Is there a correlation in between type of Injury and Implemented protection? Form hypothesis, check for conditions, and conduct a statistical test.  

```{r}
(two_way_table <- accidents %>% 
  group_by(Injury, Restraint) %>% 
  tally() %>% 
  ungroup() %>% 
  spread(Restraint, n))
```

Chi square test for independence. 

Conditions for the test:

* dataset is independent

* expected cases should be more than 5


H0: There is no correlation in between type of injury and Implemented protection in car accidents where children were passengers.

H0: There is correlation in between type of injury and Implemented protection in car accidents where children were passengers.

alpha significance level - 0.05

```{r}
colnames(two_way_table)
```


```{r}
sum_all <- sum(two_way_table$`Car Seat`) + 
  sum(two_way_table$`Lap and Shoulder Belt`) + 
  sum(two_way_table$`Lap-Only Belt`) +
  sum(two_way_table$`No Restraint`)

two_way_table %>% mutate(CS_exp = (`Car Seat` + 
                                     `Lap and Shoulder Belt` + 
                                     `Lap-Only Belt` + 
                                     `No Restraint`)*sum(two_way_table$`Car Seat`)/sum_all) %>% 
  mutate(LSB_exp = (`Car Seat` + 
                      `Lap and Shoulder Belt` + 
                      `Lap-Only Belt` + 
                      `No Restraint`)*sum(two_way_table$`Lap and Shoulder Belt`)/sum_all) %>%
  mutate(LOB_exp = (`Car Seat` + 
                      `Lap and Shoulder Belt` + 
                      `Lap-Only Belt` + 
                      `No Restraint`)*sum(two_way_table$`Lap-Only Belt`)/sum_all) %>%
  mutate(NR_exp = (`Car Seat` + 
                     `Lap and Shoulder Belt` + 
                     `Lap-Only Belt` + 
                     `No Restraint`)*sum(two_way_table$`No Restraint`)/sum_all)
  
```

All expected values are above 5.

* short version

```{r}
two_way_table %>% select(-1) %>% 
chisq.test()
```

We reject null hypothesis in favour of the alternative. There is correlation in between type of injury and Implemented protection in car accidents where children were passengers.


* long version

```{r}
two_way_table <- two_way_table %>% 
  mutate(CS_exp = (`Car Seat` + 
                     `Lap and Shoulder Belt` + 
                     `Lap-Only Belt` + 
                     `No Restraint`)*sum(two_way_table$`Car Seat`)/sum_all) %>% 
  mutate(LSB_exp = (`Car Seat` + 
                      `Lap and Shoulder Belt` + 
                      `Lap-Only Belt` + 
                      `No Restraint`)*sum(two_way_table$`Lap and Shoulder Belt`)/sum_all) %>%
  mutate(LOB_exp = (`Car Seat` + 
                      `Lap and Shoulder Belt` + 
                      `Lap-Only Belt` + 
                      `No Restraint`)*sum(two_way_table$`Lap-Only Belt`)/sum_all) %>%
  mutate(NR_exp = (`Car Seat` + 
                     `Lap and Shoulder Belt` + 
                     `Lap-Only Belt` + 
                     `No Restraint`)*sum(two_way_table$`No Restraint`)/sum_all)

```



```{r}
(chi2_stat <- sum(((two_way_table$`Car Seat` - two_way_table$CS_exp)/ two_way_table$CS_exp^0.5)^2) +
   sum(((two_way_table$`Lap and Shoulder Belt` - two_way_table$LSB_exp)/ two_way_table$LSB_exp^0.5)^2) +
   sum(((two_way_table$`Lap-Only Belt` - two_way_table$LOB_exp)/ two_way_table$LOB_exp^0.5)^2) +
   sum(((two_way_table$`No Restraint` - two_way_table$NR_exp)/ two_way_table$NR_exp^0.5)^2))
```

```{r}
(dof <- 4*3)
```

```{r}
ggplot(data.frame(x = seq(0, 25, length=100)), aes(x = x)) +
  stat_function(fun = dchisq, args = list(df = dof), color = 'blue') + 
  geom_vline(aes(xintercept = chi2_stat),  color = 'red')
```

```{r}
(p_value <- 1 - pchisq(chi2_stat, df = dof))
```

We reject null hypothesis in favour of the alternative. There is correlation in between type of injury and Implemented protection in car accidents where children were passengers.


\pagebreak

# 3.Wild blueberries yield prediction




Three datasets about wild blueberry farming are provided: 

* *data/blueberries_insects.csv* contains information about pollinating insects presence

* *data/blueberries_weather.csv* contains information about weather (temperature and rain)

* *data/blueberries_yield.csv* contains information about size of the fruit, seeds, and final yield. 

a) Join all three datasets.

```{r}
blue_insects <- readr::read_csv("data/blueberries_insects.csv")
blue_weather <- readr::read_csv("data/blueberries_weather.csv")
blue_size <- readr::read_csv("data/blueberries_yield.csv")
```

```{r}
blue <- blue_insects %>% left_join(blue_size, by = join_by(plotID)) %>% left_join(blue_weather, by = join_by(plotID))
```

b) Which variables have statistically significant influence on the blueberry yield? Create multiple regression model and tune it.


```{r}
fit <- lm(yield ~ clonesize+
            honeybee+
            bumbles+
            andrena+
            osmia+              
            MaxTemp+ 
            MinTemp+ 
            AverageTemp+
            # RainingDays+ 1st and only removed      
            AverageRainingDays+
            fruitset+
            fruitmass+        
            seeds, 
          data = blue)
```



```{r}
summary(fit)
```


yield = -12670.617 + clonesize* (-8.068) + honeybee * 50.382 + bumbles * 232.006 + andrena * 354.130 + osmia * 557.956 + MaxTemp * -3357.676 + 
MinTemp * -4052.480 + AverageTemp * 7213.292 + AverageRainingDays * -937.701 + fruitset * 8835.374 + fruitmass * -26593.008 + seeds * 349.116


Following variables have statistically significant influence on the bluberry yield:
clonesize, honeybee, bumbles, andrena, osmia, MaxTemp, MinTemp, AverageTemp, AverageRainingDays, fruitset, fruitmass, seeds.


c) What should be satisfied for model (3b) to be valid. Check if the model you created is valid?

* linearity

* nearly normal residuals

* constant variability

* independent observations


```{r}
blue %>% select(-1) %>% 
  gather(-yield, key = "var", value = "value") %>% 
  ggplot(aes(x = value, y = yield, )) +
    geom_point() +
    facet_wrap(~ var, scales = "free") +
    theme_bw()
```

All numerical variables, with continous scale have clear linear trend. 
'AverageTemp' has a more quadratic function tendency, however it's hard to pinpoint.

```{r}
plot(fit)
```

No trend visible in the first plot. There seems to be consant variability to residuals.
q-q plot also suggests that there is normal distribution to residuals. 

We assume observations are independent. 



d) Construct confidence interval for multiplication parameter of 'seeds' variable.

We are 95% confident that 'seeds' multiplier value for linear regression fit is between:

```{r}
summary(fit)$coefficients[13] - 1.96 * summary(fit)$coefficients[26]
```

and

```{r}
summary(fit)$coefficients[13] + 1.96 * summary(fit)$coefficients[26]

```


